---
title: 'ggplot2 : Introductory course'
author: "Aurélien Béliard & Gabriel Pires"
date: "11/12/2020"
output:
  pdf_document: default
  keep_tex: yes
  toc: yes

---

# Loading packages
```{r}
library(tidyverse)
library(ggplot2)
library(plotly)
library(ggpubr)
library(extrafont)

```

```{r, include=FALSE}
loadfonts(device = "win")
```


# Visualizing dataset
```{r}
mpg
```

```{r}
summary(mpg)
```


# Doplots

You can create a dot plot using the following function :
```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy))
```

Pretty basic, now let's add some colors : 

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class))
```

We can also add another layer as : 

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class))+
  geom_line(aes(x=displ, y=hwy, color=class))
```

Let's now play with the shape and size !

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class), shape="diamond", size=5)+
  geom_line(aes(x=displ, y=hwy, color=class))
```


Now that we have the plot we wanted, let's arrange it to make it prettier and easier to read.
First, let's change the names of the axes to make it clearier.

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class), shape="diamond", size=5)+
  geom_line(aes(x=displ, y=hwy, color=class))+
  xlab("Engine displacement (L)")+
  ylab("Highway miles/gallon")
```

We could also add a title.

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class), shape="diamond", size=5)+
  geom_line(aes(x=displ, y=hwy, color=class))+
  xlab("Engine displacement (L)")+
  ylab("Highway miles/gallon")+
  ggtitle("Highway miles/gallon as a function of Engine displacement")
```

But that's a lot of functions to add to our ggplot object. We could also use the function \textit{{\color{blue}labs}()}.

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class), shape="diamond", size=5)+
  geom_line(aes(x=displ, y=hwy, color=class))+
  labs(x="Engine displacement (L)",
  y="Highway miles/gallon",
  title="Highway miles/gallon as a function of Engine displacement",
  color="Car classes")
```

With this function we can also replace the name of the aesthetics, color, shape, size, fill, group, etc.

# Barplots

```{r}
ggplot(mpg)+
  geom_bar(aes(x=displ, color=class))

```

As you can see, by filling only the color aesthetics, we only color the contour of the bars. To fill the bars with a color, we can use the fill aesthetic as follow : 

```{r}
ggplot(mpg)+
  geom_bar(aes(x=displ, color=class, fill=class))+
  labs(x="Engine displacement (L)", 
       color="Car classes", 
       fill="Car classes", 
       title="Engine displacement distribution")
```

The \textit{x} axis can be treated either as a continuous variable or a discrete variable. 
To treat the \textit{x} axis as a continuous variable, use the \textit{{\color{blue} as.numeric}()} function as follow : 

```{r}
mpg$displ<-as.numeric(mpg$displ)

ggplot(mpg)+
  geom_bar(aes(x=displ, color=class, fill=class))+
  labs(x="Engine displacement (L)", 
       color="Car classes", 
       fill="Car classes", 
       title="Engine displacement distribution")


```


If you consider that your variable should be discrete, then use \textit{{\color{blue} as.factor}()} function as follow : 

```{r}
mpg$displ<-as.factor(mpg$displ)

ggplot(mpg)+
  geom_bar(aes(x=displ, color=class, fill=class))+
  labs(x="Engine displacement (L)", 
       color="Car classes", 
       fill="Car classes", 
       title="Engine displacement distribution")


```


In such cases, the \textit{x} axis can be blurry and sometimes, the labels are not clear. To make it clearier, one could use the functions  \textit{{\color{blue} scale\_x\_continuous}()} or \textit{{\color{blue} scale\_x\_discrete}()} as : 


```{r}
mpg$displ<-as.numeric(mpg$displ)

ggplot(mpg)+
  geom_bar(aes(x=displ, color=class, fill=class))+
  labs(x="Engine displacement (L)", 
       color="Car classes", 
       fill="Car classes", 
       title="Engine displacement distribution")+
  scale_x_continuous(breaks=seq(0, max(mpg$displ), 5))


```


\underline{NB} : Here, I call the max function to get the maximum value displayed on the \textit{x} axis. By doing so, we suppose that the maximum value can change from a dataset to another, so it's better to never use a maximum value chosen by hand.

Of course, here it doesn't make sense to convert the engine displacement values as factors. So we will try with another variable which is more suitable for this kind of work. We will use the class variable for our discrete variable. With \textit{{\color{blue}scale\_x\_discrete}()} we can assign another name to each label as follow : 

```{r}

ggplot(mpg)+
  geom_bar(aes(x=class, fill=fl))+
  labs(x="Car classes", 
       color="Fuel type", 
       fill="Fuel type", 
       title="Car classes per fuel type")+
  scale_x_discrete(labels=c("2seater"="2SEAT", 
                            "compact"="Compact", 
                            "midsize"="Midsize", 
                            "minivan"="Minivan", 
                            "pickup"="Pickup", 
                            "subcompact"="Subcompact", 
                            "suv"="SUV"))
  

```



# Boxplot

To create a boxplot with ggplot2, we use the geom_function \textit{{\color{blue} geom\_boxplot}()} as follow :

```{r}

ggplot(mpg, aes(x=class, y=hwy))+
  geom_boxplot(aes(fill=class))+
  labs(x="Car classes", 
       color="Fuel type", 
       fill="Fuel type", 
       title="Car classes per fuel type")+
  ylim(0, 50)+
  scale_x_discrete(labels=c("2seater"="2SEAT", 
                            "compact"="Compact", 
                            "midsize"="Midsize", 
                            "minivan"="Minivan", 
                            "pickup"="Pickup", 
                            "subcompact"="Subcompact", 
                            "suv"="SUV"))
  

```

If we want the significance level between each group, we can use the \textit{{\color{blue} stat\_compare\_means}()} function as follow : 

```{r}

box<-ggplot(mpg, aes(x=class, y=hwy))+
  geom_boxplot(aes(fill=class))+
  stat_compare_means(aes(label = ..p.signif..),
                     method="wilcox.test", 
                     label.x = 1.5, 
                     label.y = 50, 
                     ref.group = ".all.")+
  labs(x="Car classes", 
       color="Car classes", 
       fill="Car classes", 
       title="Highway miles per gallon as a function of the car classes")+
  ylim(0, 50)+
  scale_x_discrete(labels=c("2seater"="2SEAT", 
                            "compact"="Compact", 
                            "midsize"="Midsize", 
                            "minivan"="Minivan", 
                            "pickup"="Pickup", 
                            "subcompact"="Subcompact", 
                            "suv"="SUV"))
  
box
```

# Stripchart

A stripchart or jitter is a plot that produces a one dimensional dot plot of the given data. We create it with the \textit{{\color{blue} geom\_jitter}()} function.

```{r}
ggplot(mpg, aes(cty, hwy, color=factor(cyl)))+
  geom_jitter()+
  labs(
    x = "City mileage/gallon",
    y = "Highway mileage/gallon",
    colour = "Cylinders",
    title = "Highway and city mileage are highly correlated"
  )

```

We can map the tendency of the dots using the \textit{{\color{blue} abline}()} function as follow : 

```{r}
jitt<-ggplot(mpg, aes(cty, hwy, color=factor(cyl)))+
  geom_jitter()+
  labs(
    x = "City mileage/gallon",
    y = "Highway mileage/gallon",
    colour = "Cylinders",
    title = "Highway and city mileage are highly correlated"
  )+
  geom_abline(colour="blue", 
              size=1)
jitt
```

So, as you can see, the line is not mapping our dots at all. To do so, it's better to perform a linear regression to estimate the slope and intercept of the line. 

```{r}

reg<-lm(hwy~cty, data=mpg)
reg

```



```{r}
jitt<-ggplot(mpg, aes(cty, hwy, color=factor(cyl)))+
  geom_jitter()+
  labs(
    x = "City mileage/gallon",
    y = "Highway mileage/gallon",
    colour = "Cylinders",
    title = "Highway and city mileage are highly correlated"
  )+
  geom_abline(colour="blue", 
              size=1, 
              intercept = reg$coefficients[[1]], 
              slope=reg$coefficients[[2]])
jitt
```

And now we have a very correlation between our \textit{x} and \textit{y} axes.

# Theme creation

There exist several themes in ggplot2, black and white, gray, void, minimal, classic, light and so on. But what if we want our own theme, with the colors we want, a grid or not, and a specific font ? 
We must create our own theme to do that. Here we will use the black and white theme as a base theme and the function \textit{{\color{blue} %+replace%}} which replaces the entire element in a theme and if any element of a theme is not specified in the new theme, it will not be present in the resulting theme.

```{r}

my_theme<-function(){
  theme_bw() %+replace%
  theme(plot.title = element_text(hjust = 0.5, 
                                  face="bold", 
                                  colour="blue", 
                                  family = "Times"), #centers the title
        panel.border=element_blank(), 
        panel.grid.major = element_blank(), #to remove the grid
        panel.grid.minor = element_blank(), #to remove the grid
        panel.background = element_rect(fill="white"), #to fill the background with a white color
        axis.line = element_blank(), 
        legend.justification = c(0, 1))
}


```

To apply the theme to your own plot, just add the newly created theme as if you were adding a new layer.

```{r}
box+my_theme()
```

```{r}
jitt+my_theme()
```

