---
title: 'ggplot2 : Introductory course'
author: "Aurélien Béliard & Gabriel Pires"
date: "11/12/2020"
output:
  pdf_document: default
  keep_tex: yes
  toc: yes

---

# Loading packages
```{r}
library(tidyverse)
library(ggplot2)
library(plotly)
```


# Visualizing dataset
```{r}
mpg
```

```{r}
summary(mpg)
```


# First plot

You can create a dot plot using the following function :
```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy))
```

Pretty basic, now let's add some colors : 

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class))
```

We can also add another layer as : 

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class))+
  geom_line(aes(x=displ, y=hwy, color=class))
```

Let's now play with the shape and size !

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class), shape="diamond", size=5)+
  geom_line(aes(x=displ, y=hwy, color=class))
```


Now that we have the plot we wanted, let's arrange it to make it prettier and easier to read.
First, let's change the names of the axes to make it clearier.

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class), shape="diamond", size=5)+
  geom_line(aes(x=displ, y=hwy, color=class))+
  xlab("Engine displacement (L)")+
  ylab("Highway miles/gallon")
```

We could also add a title.

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class), shape="diamond", size=5)+
  geom_line(aes(x=displ, y=hwy, color=class))+
  xlab("Engine displacement (L)")+
  ylab("Highway miles/gallon")+
  ggtitle("Highway miles/gallon as a function of Engine displacement")
```

But that's a lot of functions to add to our ggplot object. We could also use the function \textit{{\color{blue}labs}()}.

```{r}
ggplot(mpg)+
  geom_point(aes(x=displ, y=hwy, color=class), shape="diamond", size=5)+
  geom_line(aes(x=displ, y=hwy, color=class))+
  labs(x="Engine displacement (L)",
  y="Highway miles/gallon",
  title="Highway miles/gallon as a function of Engine displacement",
  color="Car classes")
```

With this function we can also replace the name of the aesthetics, color, shape, size, fill, group, etc.


